Transmembrane helices are also an overlooked source of major histocompatibility complex class II epitopes and evolutionary more conserved than expected by chance

### 2021-11-13 bbbq_article.tex
Remove elution studies by Schellens and Bergseng

To obtain experimental evidence that epitopes derived from TMHs are presented in MHC,
we used two sources of data of naturally occurring epitopes.  
The first type of data are two randomly picked 
peptide elution studies for MHC-I \cite{schellens2015comprehensive} 
and MHC-II \cite{bergseng2015different}.
The second type of data is obtained from , 
the Immune Epitope Database (IEDB) \cite{vita2019immune}.
All these sources contain sequences of naturally occurring epitopes. 
For each of the detected epitopes,
its possible location(s) in 
a human reference proteome,
with UniProt ID UP000005640\_9606, was mapped.
For the epitopes that were present in the proteome exactly once,
the topology of the proteins in which these epitopes were located 
was predicted using both \verb;TMHMM; \cite{krogh2001predicting} 
and \verb;PureseqTM; \cite{wang2019efficient}. 
From this topology, we determined if the epitope
overlapped with a TMH.
As a second source of naturally occurring epitopes, 
we used the IEDB \cite{vita2019immune} for epitope sequences presented in healthy humans
by T cells, or as found in MHC ligand studies.

@@

To obtain experimental confirmation that peptides stemming from TMHs 
are presented in MHC-I and MHC-II,
two peptide elution studies were reanalyzed:
For MHC-I, peptides presented \emph{in vivo} 
by the (human) haplotypes HLA-A and B 
were sequenced \cite{schellens2015comprehensive},
for MHC-II these were alleles DQ2.5, DQ2.2, and 
DQ7.5 \cite{bergseng2015different}.
Figure \ref{fig:elution} shows the percentages of epitopes derived 
from TMHs
found in the MHC-I and MHC-II elution 
studies,
for the two topology prediction tools \verb;TMHMM; \cite{krogh2001predicting} 
and \verb;PureseqTM; \cite{wang2019efficient}. 
Regardless of the prediction tool, 
at least 100 epitopes were predicted to be derived from a TMH for each condition. 
From these findings, it is robustly predicted that
epitopes derived from TMHs are presented in both MHC-I and MHC-II.
See the supplementary Table \ref{tab:elution} for the exact values.

>>>>

To obtain experimental evidence that epitopes derived from TMHs are presented in MHC,
we used the Immune Epitope Database (IEDB) \cite{vita2019immune}.
This database contains the sequences of thousands of experimentally
established epitope sequences.
For each of these epitopes,
its possible location(s) in 
a human reference proteome,
with UniProt ID UP000005640\_9606, was mapped.
For the epitopes that were present in the proteome exactly once,
the topology of the proteins in which these epitopes were located 
was predicted using \verb;TMHMM; \cite{krogh2001predicting} . 
From this topology, we determined if the epitope
overlapped with a TMH.

@@

To obtain experimental confirmation that peptides stemming from TMHs 
are presented \emph{in vivo} in MHC-I and MHC-II,
we used epitope data from the IEDB.
We queried the IEDB for all linear epitopes obtained
from MHC ligand assays in healthy humans, 
for the MHC alleles as used in this study.
From these epitopes, we kept those that are found
exactly once in the human reference proteome.
We predicted the topology of the protein each epitope
was found in, using \verb;TMHMM; \cite{krogh2001predicting},
from which we conclude if the epitope is overlapping with a TMH
for at least 1 amino acid.

From these findings, it is robustly predicted that
epitopes derived from TMHs are presented in both MHC-I and MHC-II.
See the supplementary Table \ref{tab:elution} for the exact values.
